---
title: "Homework 6: Linear Models" 
output: github_document
Name: Jasmin Martinez
Date: 12/1/24
---

## Problem 2 
```{r}
library(rvest)
library(tidyverse)

url = "https://raw.githubusercontent.com/washingtonpost/data-homicides/master/homicide-data.csv"
homicides_data = read_csv(url)
head(homicides_data)
```

### Part A: Create a city_state variable (e.g. “Baltimore, MD”), and a binary variable indicating whether the homicide is solved

```{r}
homicides_data = homicides_data %>%
  mutate(city_state = str_c(city, ", ", state)) 

homicides_data = homicides_data %>%
  mutate(solved_homicides = ifelse(disposition == "Closed by arrest", 1, 0))

homicides_data = homicides_data %>%
  filter(!(city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")))

homicides_data = homicides_data %>%
  filter(victim_race %in% c("White", "Black"))

homicides_data = homicides_data %>%
  mutate(victim_age = as.numeric(ifelse(victim_age == "unknown", NA, victim_age)))

head(homicides_data)
```

### Part B: For the city of Baltimore, MD, use the glm function to fit a logistic regression with resolved vs unresolved as the outcome and victim age, sex and race as predictors

```{r}
bmore_data = homicides_data %>%
  filter(city_state == "Baltimore, MD") %>%
  filter(!is.na(solved_homicides), !is.na(victim_age), !is.na(victim_sex), !is.na(victim_race))

fit_logistic = bmore_data |> 
  glm(solved_homicides ~ victim_age + victim_sex + victim_race, data = _, family = binomial()) 

fit_logistic |> 
  broom::tidy(conf.int = TRUE) |> 
  mutate(OR = exp(estimate),
         lower_CI = exp(conf.low),
         upper_CI = exp(conf.high)) |>
  select(term, log_OR = estimate, OR, lower_CI, upper_CI) |> 
  knitr::kable(digits = 3)
```

### Part C: Run glm for each of the cities in your dataset, and extract the adjusted odds ratio (and CI) for solving homicides comparing male victims to female victims.
```{r}
city_nest = homicides_data %>%
  filter(!is.na(solved_homicides), !is.na(victim_sex), !is.na(victim_race)) %>%
  group_by(city_state) %>%
  nest() %>%
  mutate(
    model = map(data, ~ glm(solved_homicides ~  victim_sex, 
                            data = .x, family = binomial())),
    results = map(model, ~ broom::tidy(.x, conf.int = TRUE))
  ) %>%
  unnest(results) %>% 
  mutate(
    OR = exp(estimate), 
    lower_CI = exp(conf.low),   
    upper_CI = exp(conf.high), 
    log_OR = estimate   
  ) %>% 
   filter(
    term != "(Intercept)",        
    term != "victim_sexUnknown"     
  ) %>%
  select(city_state, term, log_OR, OR, lower_CI, upper_CI) 
city_nest %>%
  knitr::kable(digits = 3)
```

### Part D: Create a plot that shows the estimated ORs and CIs for each city. Organize cities according to estimated OR, and comment on the plot. 

```{r}
library(ggplot2)
library(dplyr)

city_nest %>% 
  ggplot(aes(x = reorder(city_state, OR, FUN = median, descending = TRUE), y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI, width = 0.2)) +  
  facet_wrap(~term) +
  theme(axis.text.x = element_text(angle = 80, hjust = 1)) +
  labs(
    x = "City, State of Homicides", 
    y = "Odds Ratio for Solving Homicides (M vs. F)"
  )

```

**Comment on the plot**: New York, NY is the city with the lowest odds (0.242) of solving a homicide when the victim is males compared to the victim being female. Albuquerque, NM (1.647) is the city with the highest odds of solving a homicide when the victim is males compared to the victim being female.In general, there is no explicit geographic pattern of the odds of solving a homicide when the victim is males compared to females. 

## Problem 3

### Part A: Load and clean the data for regression analysis (i.e. convert numeric to factor where appropriate, check for missing data, etc.).
```{r}
library(dplyr)
library(tidyr)

birthweight = read_csv("birthweight.csv",show_col_types = FALSE)

birthweight$babysex = factor(birthweight$babysex, levels = c(1, 2), labels = c("Male", "Female"))
birthweight$frace = factor(birthweight$frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown"))
birthweight$mrace = factor(birthweight$mrace, levels = c(1, 2, 3, 4, 8), labels = c("White", "Black", "Asian", "Puerto Rican", "Other"))

birthweight = birthweight %>%
  mutate(
    bhead = as.numeric(bhead),
    blength = as.numeric(blength),
    bwt = as.numeric(bwt),
    delwt = as.numeric(delwt),
    fincome = as.numeric(fincome),
   gaweeks = as.numeric(gaweeks),
    menarche = as.numeric(menarche),
    mheight = as.numeric(mheight),
    momage = as.numeric(momage),
    parity = as.numeric(parity),
    pnumlbw = as.numeric(pnumlbw),
    pnumsga = as.numeric(pnumsga),
    ppbmi = as.numeric(ppbmi),
    ppwt = as.numeric(ppwt),
    smoken = as.numeric(smoken),
    wtgain = as.numeric(wtgain)
  )
```
### Part B: Propose a regression model for birthweight.
```{r}
my_model = lm(bwt ~ fincome + momage + wtgain + ppbmi, data = birthweight)

my_model %>%
  broom::tidy() %>%
  knitr::kable(digits = 3)
```
#### Part B.2: Describe your modeling process and show a plot of model residuals against fitted values – use add_predictions and add_residuals in making this plot.
```{r}
birthweight_predict = birthweight %>%
  modelr::add_predictions(my_model) %>%
  modelr::add_residuals(my_model)

ggplot(birthweight_predict, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE)  +
  theme_minimal()
```

**Modeling Process**: In this model, I have included factors hypothesized to influence birthweight. I ijncluded family monthly income, mom's age at delivery, mother's weight gain during pregnancy, and mother's pre-pregnancy BMI. I decided to use this factors as they were show the effect of the mother's characteristics on the birthweight on her child. I also assessed the model's performance with a residuals vs fitted values plot, which showed that the model is appropriate.

### Part C: Compare your model to two others

#### Part C.2: One using length at birth and gestational age as predictors (main effects only)

#### Part C.3: One using head circumference, length, sex, and all interactions (including the three-way interaction) between these
